---
title: "H1"
output: html_document
date: '2022-04-10'
---

```{r include=FALSE}
#import libraries
install.packages("dplyr")
install.packages("tidyr")
install.packages("matrixcalc")
install.packages("zoo")
library(dplyr)
library(tidyr)
library(matrixcalc)
library(zoo)

```
Power consumption in USA:

The exercise deals with data from power stations in the United States on several contracts.

info cell: Reading the table Data

```{r}
#Reading Data
A <- read.delim("table.tsv")
A$DateTime <- as.POSIXct( A$megawatthours, tz = "EST", "%H:%M EST %m/%d/%Y" )

```
Data clearing:

1.Remove duplicate rows in the data frame

2.Complete missing values according to Last Observation in data frame

```{r}
A<-distinct(A) #1
na.locf(na.locf(A),fromLast=TRUE) #2

```
Cube construction:

The cube is bulid from dimension of time by days and the dimension of stations.

The measurable variable is Net generation
```{r}
# The dimensions of the cube are the Power stations and the dimension of date time.
data_cube<- with(A,cbind(BPAT=Net.generation,CISO	=Net.generation.1, CPLE=Net.generation.2,ERCO= Net.generation.3, FPL=Net.generation.4,ISNE= Net.generation.5,MISO= Net.generation.6,
                         NYIS =Net.generation.7 , PACW=Net.generation.8 , PJM=Net.generation.9 ,United_States_Lower = Net.generation.10,Date=as.integer(format(as.Date(A[,'DateTime']), "%Y%m%d"))))
```
Slicing:

Performing Slicing on Data cube, from all days to 07-14 Feb only.

```{r}
index<-which(data_cube[,'Date']>"20210206" & data_cube[,'Date']<"20210215" )
Silce_cube<-data_cube[index,]
```
Roll up:

Performing roll up - from a days and hours to days only.

```{r}
temp<-data_cube[index,]
rownames(temp) <- c(temp[,"Date"])
temp<-temp[,colnames(temp)!="Date"]
roll_up <-  rowsum(temp,row.names(temp))
data_mean<-rowMeans(roll_up[,1:11]) #Daily average performance
```
Plot:

Axis x - days

Axis y- Net generation

A graph that shows the daily average of the Net generation in the US


```{r}
data_frame<-tibble(data_mean)
days<- as.integer(substr(names(roll_up[,1]), nchar(names(roll_up[,1])) - 2 + 1, nchar(names(roll_up[,1]))) )
# create an empty plot
plot(1, type="n", xlab="", ylab="", xlim = c(days[1],days[8]), ylim=c(1505748 ,1900000 ))
# rearrange in a new, temporary dataframe
DF <- data.frame ( Time = days, Demand = data_frame)
# plot
lines( DF, col = 'darkorchid1', type = 'b' )
title(main = "Mean daily power generation in the US",
      xlab = "Days (07-14 Feb)", ylab = "Net generation",
      cex.main = 2,   font.main = 2, col.main = "darkorchid1")
```


```{r}
# The dimensions of the cube are the Power stations and the dimension of date time.
data_cube_q2<- with(A,cbind( CPLE=Net.generation.2, FPL=Net.generation.4,ISNE= Net.generation.5,
                         NYIS =Net.generation.7 , PJM=Net.generation.9 ,Time=as.integer(format(A[,'DateTime'], format = "%H"))))
data_cube_q2<-na.locf(data_cube_q2,fromLast=TRUE)
index_2<-which(data_cube_q2[,'Time']>9 & data_cube_q2[,'Time']<19 )
Silce_cube_2<-data_cube_q2[index_2,]


```


```{r}
temp_2<-data_cube_q2[index_2,]
rownames(temp_2) <- c(temp_2[,"Time"])
temp_2<-temp_2[,colnames(temp_2)!="Time"]
roll_up_2 <-  rowsum(temp_2,row.names(temp_2))


```


```{r}
# linear fit
LM <- list(  CPLE = NA,  FPL = NA,  ISNE = NA,  NYIS = NA, PJM = NA)
Lm_data<-data.frame (roll_up_2)
min_max_norm <- function(x) {
    (x - min(x)) / (max(x) - min(x))
}



mat1_data <- c(rep(c(1),9),c(10:18))
mat1 <- matrix(mat1_data,nrow=2,ncol=9,byrow=TRUE)
Demand_norm <- as.data.frame(lapply(Lm_data, min_max_norm))



"%^%" <- function(mat, power) 
  with(eigen(mat), vectors %*% (values^power * t(vectors))) 

# create an empty plot
plot(1, type="n", xlab="", ylab="",xaxt="n",xlim = c(10,18), ylim=c(0 ,1))
axis(1, at=c(10:18),labels=c("10:00","11:00","12:00","13:00","14:00","15:00","16:00","17:00","18:00"))
demands <- seq(5)
a<-0
b<-0
for ( i in demands ) {
   # rearrange in a new, temporary dataframe
   DF <- data.frame ( Time =c(10:18), Demand = Demand_norm[,i] )
   # plot
   lines( DF, col = i, type = 'b' )
   # linear fit
   X_T_X<-((mat1)%*% t(mat1))
   Vector_df1<-c(t(DF[2]))
   X_T_y<-((mat1))%*% (Vector_df1)
   LM[[ i ]] <-(X_T_X%^% (-1))%*% X_T_y
   a <- a+(LM[[ i ]])[1]
   b <- b+(LM[[ i ]])[2]
   #abline(a, b, col = i, lw =2)
}

abline(a/5, b/5, col = i, lw =2)
title(main = "Minute power demand in the east coast",
      xlab = "Time", ylab = "demand",
      cex.main = 2,   font.main = 1, col.main = "black")
legend( 'bottomleft', col = demands, pch = 19,legend = names(Demand_norm) )
```


```{r}
index_3<-which(data_cube_q2[,'Time']>19 | data_cube_q2[,'Time']<4 )
Silce_cube_3<-data_cube_q2[index_3,]
```


```{r}
temp_3<-data_cube_q2[index_3,]
rownames(temp_3) <- c(temp_3[,"Time"])
temp_3<-temp_3[,colnames(temp_3)!="Time"]
roll_up_3 <-  rowsum(temp_3,row.names(temp_3))
```


```{r}

# linear fit
LM <- list(  CPLE = NA,  FPL = NA,  ISNE = NA,  NYIS = NA, PJM = NA)
Lm_data_3<-data.frame (roll_up_3)
min_max_norm <- function(x) {
    (x - min(x)) / (max(x) - min(x))
}



mat2_data <- c(rep(c(1),8),c(1:8))
mat2 <- matrix(mat2_data,nrow=2,ncol=8,byrow=TRUE)
Demand_norm_2 <- as.data.frame(lapply(Lm_data_3, min_max_norm))

dd<-Demand_norm_2[c(4,5,6,7,1,2,3,8),]

"%^%" <- function(mat, power) 
  with(eigen(mat), vectors %*% (values^power * t(vectors))) 


# create an empty plot
plot(1, type="n", xlab="", ylab="",xaxt="n", xlim = c(1,8), ylim=c(0 ,1))
axis(1, at=c(1:8),labels=c("20:00","21:00","22:00","23:00","00:00","01:00","02:00","03:00"))
demands <-seq(5)# c(4,5,6,7,1,2,3,8)
a<-0
b<-0

for ( i in demands ) {
   # rearrange in a new, temporary dataframe
   DF <- data.frame ( Time =c(1,8), Demand = dd[,i] )
   # plot
   lines( DF[2], col = i, type = 'b' )
   # linear fit
   X_T_X<-((mat2)%*% t(mat2))
   Vector_df1<-c(t(DF[2]))
   X_T_y<-((mat2))%*% (Vector_df1)
   LM[[ i ]] <-(X_T_X%^% (-1))%*% X_T_y
   a <- a+(LM[[ i ]])[1]
   b <- b+(LM[[ i ]])[2]
   #abline(a, b, col = i, lw =2)
}
abline(a/5, b/5, col = 'gray7', lw =2)
title(main = "Minute power demand in the east coast",
      xlab = "Time", ylab = "demand",
      cex.main = 2,   font.main = 1, col.main = "black")
#legend( 'bottomleft', col = demands, pch = 19,legend = names(Demand_norm) )
```

