---
title: "H1"
output: html_document
date: '2022-04-10'
Naama Shenberger ID:211983747
Ella Hanzin ID:212028971
---
```{r include=FALSE}
#import libraries
install.packages("dplyr")
install.packages("tidyr")
install.packages("matrixcalc")
install.packages("zoo")
library(dplyr)
library(tidyr)
library(matrixcalc)
library(zoo)

```
Power consumption in USA:

The exercise deals with data from power stations in the United States on several contracts.

info cell: Reading the table Data
```{r}
#Reading Data
A <- read.delim("table.tsv")
A$DateTime <- as.POSIXct( A$megawatthours, tz = "EST", "%H:%M EST %m/%d/%Y" )

```

Data clearing:

Remove duplicate rows in the data frame

```{r}
A<-distinct(A)
```

Question 1:

Cube construction:

The cube is bulid from dimension of time by days and the dimension of stations.

The measurable variable is Net generation

```{r}
# The dimensions of the cube are the Power stations and the dimension of date time.
data_cube<- with(A,cbind(BPAT=Net.generation,CISO	=Net.generation.1, CPLE=Net.generation.2,ERCO= Net.generation.3, FPL=Net.generation.4,ISNE= Net.generation.5,MISO= Net.generation.6,
                         NYIS =Net.generation.7 , PACW=Net.generation.8 , PJM=Net.generation.9 ,United_States_Lower = Net.generation.10,Date=as.integer(format(as.Date(A[,'DateTime']), "%Y%m%d"))))

```

adjusting data cube  to time zones:

That the times of the data are in the timezon of EST and stations: BPAT, CISO, PACW are in PDT time which is 3 hours back from EST and stations: ERCO, MISO are in CDT time which is one hour back from EST, the other stations are in EDT which is equal to EST.
BPAT, CISO, PACW are in columns 1,2,9
 3 hours back=>
 row range is 4:n
 data_cube[4:n, col_number]
ERCO, MISO are in columns 4,7
 1 hour back=>
 row range is 2:(n-2)
 data_cube[2:(n-2), col_number]

EST and stations dont need the time zone to change
 row range is 1:(n-3)
 data_cube[1:(n-3), col_number]
 
datacube_timezon has n-3 rows

```{r}
n<-dim(data_cube)[1]#The number of rows in the data frame is
datacube_timezon<-cbind(data_cube[4:n,c(1,2)],data_cube[1:(n-3),3],data_cube[2:(n-2),4],data_cube[1:(n-3),c(5,6)],data_cube[2:(n-2),7],data_cube[1:(n-3),8],data_cube[4:n,9],data_cube[1:(n-3),c(10,11,12)] )
```

Slicing:

Performing Slicing on Data cube, from all days to 07-14 Feb only.

```{r}
index<-which(data_cube[,'Date']>"20210206" & data_cube[,'Date']<"20210215" )
Silce_cube<-datacube_timezon[index,]
```
Roll up:

Performing roll up - from a days and hours to days only.
```{r}
temp<-data_cube[index,]
rownames(temp) <- c(temp[,"Date"])
temp<-temp[,colnames(temp)!="Date"]
roll_up <-  rowsum(temp,row.names(temp))
data_mean<-rowMeans(roll_up[,1:11]) #Daily average performance
```

Plot:

Axis x - days

Axis y- Net generation

A graph that shows the daily average of the Net generation in the US


```{r}
data_frame<-tibble(data_mean)
days<- as.integer(substr(names(roll_up[,1]), nchar(names(roll_up[,1])) - 2 + 1, nchar(names(roll_up[,1]))) )

pdf(file = "outputHW1/Week3_Mean_daily_power.pdf",   
    width = 8, # The width of the plot in inches
    height = 5) # The height of the plot in inches


# create an empty plot
plot(1, type="n", xlab="", ylab="", xlim = c(days[1],days[8]), ylim=c(1600000 ,1900000 ))
# rearrange in a new, temporary dataframe
DF <- data.frame ( Time = days, Demand = data_frame)
# plot
lines( DF, col = 'darkorchid1', type = 'b' )
title(main = "Mean daily power generation in the US",
      xlab = "Days (07-14 Feb)", ylab = "Net generation",
      cex.main = 2,   font.main = 2, col.main = "darkorchid1")
dev.off()
```

Question 2:
Note: The variables with a lower hyphen 2_1 refer to question 2 graph number 1, because we created 2 graphs in question 2.

```{r}
# The dimensions of the cube are the Demand Power of the stations and the dimension of date time.
data_cube_2_1<- with(A,cbind( CPLE=Demand.2, FPL=Demand.4,ISNE= Demand.5,
                         NYIS =Demand.7 , PJM=Demand.9 ,Time=as.integer(format(A[,'DateTime'], format = "%H"))))

# Complete missing values according to Last Observation in data cube
data_cube_2_1<-na.locf(data_cube_2_1,fromLast=TRUE)
```

Slicing:

Performing Slicing on Data cube, from all Times in the day to 10:00-18:00.

```{r}
#Silce cube
index_2_1<-which(data_cube_2_1[,'Time']>9 & data_cube_2_1[,'Time']<19 )
Silce_cube_2_1<-data_cube_2_1[index_2_1,]
```


Grouping:

Grouping the Data cube by time groups

```{r}
temp_2_1<-data_cube_2_1[index_2_1,]
rownames(temp_2_1) <- c(temp_2_1[,"Time"])
temp_2_1<-temp_2_1[,colnames(temp_2_1)!="Time"]
roll_up_2_1 <-  rowsum(temp_2_1,row.names(temp_2_1))
```

Perform a linear regression to get the power requirement in minutes.
We calculated the average regression lines of five stations on the east coast.

```{r}
# linear fit
LM <- list(  CPLE = NA,  FPL = NA,  ISNE = NA,  NYIS = NA, PJM = NA)
# Creating a dataframe for performing linear regression
df_2_1<-data.frame (roll_up_2_1)
# Data normalization function min-max method.
min_max_norm <- function(x) {
    (x - min(x)) / (max(x) - min(x))
}
#Data normalization
Demand_norm_2_1 <- as.data.frame(lapply(df_2_1, min_max_norm))


#Linear regression formula (U^T*U)^-1*U^T*y
#Create a U matrix for linear regression 
mat2_1_data <- c(rep(c(1),9),c(10:18))
mat2_1 <- matrix(mat2_1_data,nrow=2,ncol=9,byrow=TRUE)

#Create a power function
"%^%" <- function(mat, power) 
  with(eigen(mat), vectors %*% (values^power * t(vectors))) 

pdf(file = "outputHW1/Week3_Minute_power_demand_1.pdf",   
    width = 8, # The width of the plot in inches
    height = 5) # The height of the plot in inches
# create an empty plot
plot(1, type="n", xlab="", ylab="",xaxt="n",xlim = c(10,18), ylim=c(0 ,1))
axis(1, at=c(10:18),labels=c("10:00","11:00","12:00","13:00","14:00","15:00","16:00","17:00","18:00"))
demands <- seq(5) 
#initialization a&b
a<-0
b<-0
colors<-c("aquamarine2","cornflowerblue","coral1","blueviolet","mediumvioletred") #Vector of colors
#A loop that draws a line for each station
for ( i in demands ) {
   # rearrange in a new, temporary dataframe
   DF <- data.frame ( Time =c(10:18), Demand = Demand_norm_2_1[,i] )
   # plot
   lines( DF, col = colors[i], type = 'b' )
   # linear fit
   #Note: 'U' have been in Transformation form
   X_T_X<-((mat2_1)%*% t(mat2_1)) #U^T*U
   Vector_df1<-c(t(DF[2])) #transfer Y df to Vector
   X_T_y<-((mat2_1))%*% (Vector_df1) #U^T*y
   LM[[ i ]] <-(X_T_X%^% (-1))%*% X_T_y #(U^T*U)^-1*U^T*y
   a <- a+(LM[[ i ]])[1]
   b <- b+(LM[[ i ]])[2]
}
#demand=a+b*time
abline(a/5, b/5, col = "black", lw =2) #calculated the average regression lines of five stations on the east coast.
title(main = "Minute power demand in the east coast",
      xlab = "Time", ylab = "demand",
      cex.main = 2,   font.main = 1, col.main = "black")
legend( 'bottomleft', col = colors, pch = 19,legend = names(Demand_norm_2_1) )

dev.off()
```

Slicing:

Performing Slicing on Data cube, from all Times in the day to 20:00 to 3:00 o'clock.

```{r}
index_2_2<-which(data_cube_2_1[,'Time']>19 | data_cube_2_1[,'Time']<4 )
Silce_cube_2_2<-data_cube_2_1[index_2_2,]
```

Grouping:

Grouping the Data cube by time groups

```{r}
temp_2_2<-data_cube_2_1[index_2_2,]
rownames(temp_2_2) <- c(temp_2_2[,"Time"])
temp_2_2<-temp_2_2[,colnames(temp_2_2)!="Time"]
roll_up_2_2 <-  rowsum(temp_2_2,row.names(temp_2_2))
```


```{r}
# linear fit
LM <- list(  CPLE = NA,  FPL = NA,  ISNE = NA,  NYIS = NA, PJM = NA)
df_2_2<-data.frame (roll_up_2_2)
#Data normalization
Demand_norm_2_2 <- as.data.frame(lapply(df_2_2, min_max_norm))



#Linear regression formula (U^T*U)^-1*U^T*y
#Create a U matrix for linear regression 
mat2_2_data <- c(rep(c(1),8),c(1:8))
mat2_2 <- matrix(mat2_2_data,nrow=2,ncol=8,byrow=TRUE)

#Enter the data in the order of the hours from 20:00-03:00
order_data<-Demand_norm_2_2[c(4,5,6,7,1,2,3,8),]


pdf(file = "outputHW1/Week3_Minute_power_demand_2.pdf",   
    width = 8, # The width of the plot in inches
    height = 5) # The height of the plot in inches

# create an empty plot
plot(1, type="n", xlab="", ylab="",xaxt="n", xlim = c(1,8), ylim=c(0 ,1))
axis(1, at=c(1:8),labels=c("20:00","21:00","22:00","23:00","00:00","01:00","02:00","03:00"))
demands <-seq(5)
#initialization a&b
a<-0
b<-0
colors<-c("aquamarine2","cornflowerblue","coral1","purple3","mediumvioletred") #Vector of colors
#A loop that draws a line for each station
for ( i in demands ) {
   # rearrange in a new, temporary dataframe
   DF <- data.frame ( Time =c(1,8), Demand = order_data[,i] )
   # plot
   lines( DF[2], col = colors[i], type = 'b' )
   # linear fit
   #Note: 'U' have been in Transformation form
   X_T_X<-((mat2_2)%*% t(mat2_2)) #U^T*U
   Vector_df2<-c(t(DF[2])) #transfer Y df to Vector
   X_T_y<-((mat2_2))%*% (Vector_df2) #U^T*y
   LM[[ i ]] <-(X_T_X%^% (-1))%*% X_T_y #(U^T*U)^-1*U^T*y
   a <- a+(LM[[ i ]])[1]
   b <- b+(LM[[ i ]])[2]
}
abline(a/5, b/5, col = 'gray7', lw =2) #calculated the average regression lines of five stations on the east coast.
title(main = "Minute power demand in the east coast",
      xlab = "Time", ylab = "demand",
      cex.main = 2,   font.main = 1, col.main = "black")
legend( 'bottomleft', col = colors, pch = 19,legend = names(order_data) )

dev.off()
```

Graph conclusions:

plot Week3_Mean_daily_power 

A graph showing the daily average of electricity consumption on 07/02/21-14/02/21
It can be seen that the highest power consumption was on 12/02/2021

plot Week3_Minute_power_demand_1 

You can see that all five station lines go up towards evening hours (16:00-18:00). In the evening there is more electricity demand.

plot Week3_Minute_power_demand_2 

You can see that in the late evening (20:00-22:00) there is a high electricity demand and as you get to late at night (before dawn) the electricity demand decreases because during these hours people sleep so there is no demand for electricity.





